### 2021-08-30 publications/dora_outlier_detection_iaai.tex

\def\year{2021}\relax
%File: formatting-instructions-latex-2021.tex
%release 2021.2
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai21}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
%\nocopyright
%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
\pdfinfo{
/Title (AAAI Press Formatting Instructions for Authors Using LaTeX -- A Guide)
/Author (AAAI Press Staff, Pater Patel Schneider, Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen, Francisco Cruz, Marc Pujol-Gonzalez)
/TemplateVersion (2021.2)
} %Leave this
% /Title ()
% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
% Leave the space between \Title and the beginning parenthesis alone
% /Author ()
% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case.
% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands,
% remove them.

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.

% The file aaai21.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%

% Title

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\iffalse
\title{AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide }
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,  \\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz,
    Marc Pujol-Gonzalez
    \\
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    %If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    %For example,

    % Sunil Issar, \textsuperscript{\rm 2}
    % J. Scott Penberthy, \textsuperscript{\rm 3}
    % George Ferguson,\textsuperscript{\rm 4}
    % Hans Guesgen, \textsuperscript{\rm 5}.
    % Note that the comma should be placed BEFORE the superscript for optimum readability

    2275 East Bayshore Road, Suite 160\\
    Palo Alto, California 94303\\
    % email address must be in roman text type, not monospace or sans serif
    publications21@aaai.org

    % See more examples next
}
\fi
\iffalse
%Example, Single Author, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
\title{My Publication Title --- Single Author}
\author {
    % Author
    Author Name \\
}

\affiliations{
    Affiliation \\
    Affiliation Line 2 \\
    name@example.com
}
\fi

%Example, Multiple Authors, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
\title{Domain-agnostic Outlier Ranking Algorithms (DORA)---A Configurable Pipeline for Facilitating Outlier Detection in Scientific Datasets}
\author {
    % Authors
    First Author Name,\textsuperscript{\rm 1}
    Second Author Name, \textsuperscript{\rm 2}
    Third Author Name \textsuperscript{\rm 1} \\
}
\affiliations {
    % Affiliations
    \textsuperscript{\rm 1} Affiliation 1 \\
    \textsuperscript{\rm 2} Affiliation 2 \\
    firstAuthor@affiliation1.com, secondAuthor@affilation2.com, thirdAuthor@affiliation1.com
}

\begin{document}

\maketitle

\begin{abstract}
Automatic detection of out-of-distribution (OOD) samples or features is
universally needed when working with scientific datasets. For example,
OOD detection methods can be used for cleaning datasets or flagging
novel samples in order to guide instrument acquisition or scientific analysis.
We present Domain-agnostic Outlier Ranking Algorithms (DORA) to provide
a configurable pipeline that makes it easy for scientists to apply and evaluate
outlier detection methods in a variety of domains. DORA allows users to 
configure outlier detection experiments by specifying the location of their 
dataset(s), the input data type, feature extraction methods, and which 
algorithms should be applied. DORA supports image, raster, time series, 
or feature vector input data types and outlier detection methods that include
 Isolation Forest, Discovery via Eigenbasis Modeling of Uninteresting Data 
 (DEMUD), principal component analysis (PCA), Reed-Xiaoli (RX) detector, 
 Local RX, negative sampling, and probabilistic autoencoder. Given a set of
  data samples, each algorithm assigns an outlier score to each sample. DORA
   provides results organization and visualization modules to help users 
   process the results, including sorting samples by outlier score, evaluating 
   outlier recall for a set of known/labeled outliers, clustering groups of 
   similar outliers together, and causal graphs. To demonstrate the 
   cross-domain applicability of DORA, we conducted outlier detection
    experiments for four use cases: 1) cleaning noisy ground-truth datasets
     (Earth Science), 2) detecting volcanic thermal anomalies (Earth Science), 
     3) flagging non-astrophysical objects in the Dark Energy Survey 
     (astrophysics), and 4) novelty-guided onboard targeting of instruments
      for Mars rovers (planetary science). [We showed that ...]
\end{abstract}

\section{Introduction}
\begin{itemize}
\item Framing of outlier detection as a domain-agnostic problem that is well-
suited to cross-disciplinary applications
\item Example applications of outlier detection across science application 
domains
\item Challenges/barriers to entry for domain scientists who could benefit from
 outlier detection 
\item DORA as a solution to lower these barriers to entry and help integrate 
outlier detection methods into existing scientific workflows across variety of 
domains
\end{itemize}

\section{Related Work}
\begin{itemize}
\item Methods for outlier detection
\item Applications of outlier detection related to our focus domains 
(e.g., \citet{kerner2020comparison})
\item Tools for applying outlier detection or other methods across domains?

\end{itemize}

\section{Methods}

\subsection{DORA configurable pipeline} Steven please add details about the 
approach/design choices for this pipeline. Mention that additional data types, 
feature extraction methods, ranking algorithms, and results organization and 
visualization modules can easily be added and integrated if they follow the API. 
Hannah to add architecture figure. Add descriptions of each layer of modules: 
\paragraph{Data types} Hannah
\paragraph{Feature extraction} Hannah 
\paragraph{Ranking algorithms} We implemented 7 algorithms for scoring and 
ranking samples by outlierness. [Describe each algorithm briefly and how they 
complement each other.]
\paragraph{Results organization} Eric, Umaa, Steven, Vinay
\paragraph{Outlier visualization} Jake?

\section{Experiments}

\subsection{Datasets}
Describe the datasets overall and our motivation for choosing these. Maybe add
 a table with number of samples, number of reference known outliers (if 
 available), etc, if room?

\subsubsection{Astrophysical objects in Dark Energy Survey}
Umaa

\subsubsection{Targets in Mars rover images}
Kiri

\subsubsection{Satellite time series for ground-truth observations}
Hannah: Describe the Earth crop field data set and applications

\subsubsection{Volcanic thermal anomalies}

\subsubsection{MNIST and FashionMNIST}
We added this to have a traditional benchmark and use case demonstration.

\subsection{Results}
Discuss the overall approach to the setup of the experiments. In each 
subsection discuss specific setup details (including evaluation metrics) for 
each dataset then the results from the experiments.

\subsubsection{Astrophysical objects in Dark Energy Survey}
Umaa

\subsubsection{Targets in Mars rover images}
Kiri

\subsubsection{Satellite time series for ground-truth observations}
Hannah: Describe the Earth crop field data set and applications

\subsubsection{Volcanic thermal anomalies}

\subsubsection{MNIST and FashionMNIST}
We added this to have a traditional benchmark and use case demonstration.

\section{Discussion}
\begin{itemize}
\item Discuss results from experiments
\item Challenges of evaluating results in outlier detection?
\item Downstream use of results in science applications/Path to deployment
\end{itemize}

\section{Conclusions}
Summarize conclusions and plans for future work, including public 
pip-installable package

[Note: these datasets could also be used by other researchers as OOD
 benchmark tasks?]

\bibliography{dora_references}

\section{ Acknowledgments}
Funding/support: JPL CIF, SMD etc. Acknowledge people who contributed to 
datasets e.g. FAO, Matt Pritchard, DES, other CIF team members?

\end{document}